---
title: "Longitudinal Analysis of the Brumandhino Data"
author: "Daniel Salnikov"
format:
  html:
    fig-width: 16
    fig-height: 12
    theme: cyborg
    toc: true
editor: visual
---

```{r setup}
#| echo: false
#| warning: false

Bru_data <- load("~/Documents/PhD/my_stuff/papers/communal_gnar_costa_rica/dynamic_GNAR_revised/Bru_data.RData")
require(GNAR)
require(igraph)
require(Rcpp)
require(RcppArmadillo)
require(fastCorbit)
require(glasso)
require(sparsevar)
require(ggplot2)
require(viridis)
require(sf)
require(ggplot2)
require(ggspatial)
require(prettymapr)
sourceCpp('/Users/danielsalnikov/Documents/PhD/code_scripts/fast_nacf_cpp_files/nacf_inner.cpp')
sourceCpp('/Users/danielsalnikov/Documents/PhD/code_scripts/fast_nacf_cpp_files/get_k_stage_adjacency_list.cpp')
source("/Users/danielsalnikov/Documents/PhD/code_scripts/vs_code_editing/community_GNAR_simulation")
source('/Users/danielsalnikov/Documents/PhD/code_scripts/fast_nacf_cpp_files/nacf_pnacf_rwrapper.r')
source('/Users/danielsalnikov/Documents/PhD/code_scripts/network_autocorrelation/corbit_scripts/weight_adjustment_methods.R')
source("/Users/danielsalnikov/Documents/PhD/code_scripts/communal_gnar/global_gnar_fitting.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/vs_code_editing/structural_GNAR_fitting_methods.r")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/vs_code_editing/struc_GNAR_sim_setup.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/communal_gnar/community_interactions_fitting_methods.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/vs_code_editing/one-stepcommunity_GNAR_prediction.r", encoding = "UTF-8")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/communal_gnar/community_gnar_fitting_methods.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/communal_gnar/communal_gnar_simulation_methods.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/communal_gnar/global_gnar_fitting.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/communal_gnar/com_GNARfit_tv_weights.R")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/vs_code_editing/community_GNAR_simulation")
source("/Users/danielsalnikov/Documents/PhD/code_scripts/vs_code_editing/community_GNARfit.r")
source("~/Documents/PhD/code_scripts/gnar_correlation_structure/correlation_structure_plots.R")
require(fda)
```

# Code for Reproducing Analyses in Section 6

This document contains code and explanations to reproduce the results in Section 6. Most of the source code is not printed, however, it can be easily accessed by looking at the code blocks in the `Quarto` file.

## Construction of UTM-distance \`\`Gram" Matrix

We provide the code for constructing the \`\`Gram" kernel matrix, where each entry is given by $\exp[- \{ \delta_{\mathrm{UTM}} (i, \, j) / (2 \cdot 200 )\}^2 ]$, where $\delta_{\mathrm{UTM}} (i, \, j)$ is the distance in UTM units between locations $i$ and $j$. We begin by plotting the geographical locations superimposed on a map with UTM coordinates on the $\mathrm{x}$- and $\mathrm{y}$-axis.

```{r mapPlot}
#| echo: false

# 1. Define your UTM coordinates
# Easting: 591817, Northing: 7775020
# Zone 23S (EPSG: 31983 for SIRGAS 2000 / UTM zone 23S)
coords_df <- data.frame(
  id = "Target Location",
  x = X,
  y = Y
)

# 2. Convert to an 'sf' object 
# We set the CRS to 31983, which is the standard for UTM 23S in Brazil
point_sf <- st_as_sf(coords_df, coords = c("x", "y"), crs = 31983)

# 3. Create the plot
ggplot() +
  # Add a basemap layer (requires internet connection)
  annotation_map_tile(type = "osm", zoom = 18) + 
  # Plot the point
  geom_sf(data = point_sf, color = "black", size = 4, shape = 16) +
  # Formatting
  ggtitle("Ground Displacements at the Brumadinho Dam by Location in Belo Horizonte, Brazil (UTM 23S) \n")+
  theme_minimal()+
  theme(plot.title = element_text(color="black", size=22, face="bold"),
        axis.text.x = element_text(color="black", size=18, face="bold"),
        axis.text.y = element_text(color="black", size=18, face="bold"))
```

The resulting network is obtained by fitting the `graphical_lasso` to the \`\`Gram" matrix constructed above. We plot the network below.

```{r netPlot}
#| warning: false

UTM_coords <- matrix(c(X, Y), nrow = 98, ncol =2)
UTM_dist <- function(x1, x2) {
  a = as.numeric(t(x1 - x2) %*% (x1 - x2))
  return(sqrt(a))
}
UTM_dist_matrix <- matrix(rep(0.0, 98*98), nrow = 98, ncol = 98)
#UTM_dist_matrix <- vapply(1:98, function(i){UTM_dist_inner(UTM_coords[i, ])}, rep(0.0, 98))
for (i in 1:98) {
  UTM_dist_matrix[i, ] = vapply(1:98, function(j){UTM_dist(UTM_coords[i, ], UTM_coords[j, ])}, 0.0)
}
View(UTM_dist_matrix)
mean_dist = median(UTM_dist_matrix)
UTM_kernel = exp(- (UTM_dist_matrix / (2 * 200))^2)
hist(UTM_dist_matrix[UTM_dist_matrix > 0], breaks = 50)
net_mat <- glasso(s = UTM_kernel, rho = 0.12, penalize.diagonal = FALSE)
View(net_mat$wi)
b =net_mat$wi
b[abs(b) > 0.0] = 1
diag(b) = 0
net_result = graph_from_adjacency_matrix(b, 'undirected')
V(net_result)$Xcoord <- X
V(net_result)$Ycoord <- Y
V(net_result)$color <- rep('black', 98)
node_labels <- vapply(1:98, function(i) {return(paste0('(', as.character(i), ')'))}, '')
group_legend_labels <- c(vapply(1:3, function(i) {c(paste0('Location: ', as.character(i), ' & Satellite: 1'), 
                                                    paste0('Location: ', as.character(i), ' & Satellite: 2'))}, c('', '')))
location <- matrix(c(V(net_result)$Xcoord, V(net_result)$Ycoord), ncol = 2)
plot.igraph(net_result, vertex.label='', vertex.label.cex = .5, vertex.color = 'black', edge.width = 0.5, vertex.label.color = 'black', 
            vertex.shape = 'circle', vertex.size = 500, mark.col = c('#ADD5D7', '#B3DBB6', '#FDB06C', '#7871A8', '#AACAF5', '#F6C5E2'),
            mark.groups = list(c(1:10), c(11:20), c(21:37), c(72:86), c(38:60), c(c(61:71), c(87:98))), mark.border = NA,
            layout = location, rescale=FALSE, asp = 0, xlim = range(V(net_result)$Xcoord), ylim = range(V(net_result)$Ycoord))
title(main = "\n Brumadinho Dam Location Network", cex.main=3)
legend("bottomright",
       legend = group_legend_labels,
       pch = 16,
       text.col = 'black',
       pt.cex = 2,
       cex = 2,
       bty = 'o', 
       horiz = FALSE,
       col = c('#ADD5D7', '#B3DBB6', '#FDB06C', '#7871A8', '#AACAF5', '#F6C5E2')
       )
```

## Model Order Selection and Exploratory Analysis

We begin by specifying model order for each group. There are six groups in total, one for each satellite and location pair.

```{r fitSetup}
locind_edt <- c(rep(1, 10), rep(2, 10), rep(3, 17), rep(5, 23), rep(6, 11), rep(4, 15), rep(6, 12))
bur_netGNAR <- igraphtoGNAR(net_result)
S1 <- as.matrix(GNARtoigraph(bur_netGNAR))
D = distances(graph_from_adjacency_matrix(S1))
D[!is.finite(D)] <- 0
rmax = max(D)
W_bru_data <- fastCorbit::weights_matrix_cpp(bur_netGNAR, rmax)
# Set community membership values
loc1_sat1 <- vapply(1:98, function(x) {return(ifelse(locind_edt[x]==1, 1, 0))}, 0)
loc1_sat2 <- vapply(1:98, function(x) {return(ifelse(locind_edt[x]==2, 1, 0))}, 0)
loc2_sat1 <- vapply(1:98, function(x) {return(ifelse(locind_edt[x]==3, 1, 0))}, 0)
loc2_sat2 <- vapply(1:98, function(x) {return(ifelse(locind_edt[x]==4, 1, 0))}, 0)
loc3_sat1 <- vapply(1:98, function(x) {return(ifelse(locind_edt[x]==5, 1, 0))}, 0)
loc3_sat2 <- vapply(1:98, function(x) {return(ifelse(locind_edt[x]==6, 1, 0))}, 0)

bru_data_community_indicators <- list(loc1_sat1, loc1_sat2, loc2_sat1, loc2_sat2, loc3_sat1, loc3_sat2)
W_bru_data <- fastCorbit::weights_matrix_cpp(bur_netGNAR, rmax)
#bru_data_interactions <- list(c(2), c(1), c(4), c(3), c(6), c(5))
bru_data_interactions <- list(c(2:6), c(1, 3:6), c(1,2, 4:6), c(1, 2, 3, 5, 6), c(1, 2, 3, 4, 6), c(1:5))
alpha_order <- rep(1, 6)
beta_order <- list(c(2), c(2), c(2), c(2), c(3), c(3)) # one for each community
```

We plot a histogram of the centred longitudinal data and analyse Corbit and R-Corbit plots for verifying model order selection of one-lag and stage-three $r > 1$.

```{r EDA}
bru_vts <- vapply(1:98, function(x) {return(100.0 * MXD2[, x] - colMeans(100.0 * MXD2)[x])}, rep(0.0, 44))
hist(recursive_dmat_cols_cpp(lapply(1:43, function(j) {return(as.matrix(bru_vts[2:44, j]))})), breaks = 50, xlab = 'y', main ='Bru VTS')

nacf_raw_data <- corbit_plot_cpp(MXD2, bur_netGNAR, 10, 4, W_bru_data)
max(abs(nacf_raw_data))
corbit_plot_cpp(MXD2, bur_netGNAR, 10, 4, W_bru_data, partial = "yes")


# Community-wise decomposition for each vector time series
loc1_sat1_nts <- t(vapply(seq(1:44), function(x) {return(loc1_sat1 * MXD2[x, ])}, rep(0.0, 98)))
loc1_sat2_nts <- t(vapply(seq(1:44), function(x) {return(loc1_sat2 * MXD2[x, ])}, rep(0.0, 98)))
loc2_sat1_nts <- t(vapply(seq(1:44), function(x) {return(loc2_sat1 * MXD2[x, ])}, rep(0.0, 98)))
loc2_sat2_nts <- t(vapply(seq(1:44), function(x) {return(loc2_sat2 * MXD2[x, ])}, rep(0.0, 98)))
loc3_sat1_nts <- t(vapply(seq(1:44), function(x) {return(loc3_sat1 * MXD2[x, ])}, rep(0.0, 98)))
loc3_sat2_nts <- t(vapply(seq(1:44), function(x) {return(loc3_sat2 * MXD2[x, ])}, rep(0.0, 98)))


# R-Corbit plots
# NACF
# PNACF
r_corbit_plot_cpp(list(loc1_sat1_nts, loc1_sat2_nts, loc2_sat1_nts, loc2_sat2_nts, loc3_sat1_nts, loc3_sat2_nts), list(bur_netGNAR, bur_netGNAR, bur_netGNAR, bur_netGNAR, bur_netGNAR, bur_netGNAR),
                  max_lag = 10, max_stage = 4, list(W_bru_data, W_bru_data, W_bru_data, W_bru_data, W_bru_data, W_bru_data), c("loc1_sat1", "loc1_sat2", "loc2_sat1", "loc2_sat2", "loc3_sat1", "loc3_sat2"), 
                  same_net = "no", partial = "no")

r_corbit_plot_cpp(list(loc1_sat1_nts, loc1_sat2_nts, loc2_sat1_nts, loc2_sat2_nts, loc3_sat1_nts, loc3_sat2_nts), list(bur_netGNAR, bur_netGNAR, bur_netGNAR, bur_netGNAR, bur_netGNAR, bur_netGNAR),
                  max_lag = 5, max_stage = 4, list(W_bru_data, W_bru_data, W_bru_data, W_bru_data, W_bru_data, W_bru_data), c("loc1_sat1", "loc1_sat2", "loc2_sat1", "loc2_sat2", "loc3_sat1", "loc3_sat2"), 
                  same_net = "no", partial = "yes")
```

## Fitting the Dynamic GNAR Model

We proceed to fit each of the models in the paper and compute the corresponding diagnostic statistics. We fill the table one model at a time and plot a histogram of the residuals and their evolution as a linear function of time: $\hat{u}_{i, t_j} = \hat{b}_{0, i} + \hat{b}_{1, i} \, t_j$ for $i = 1, \dots, 98$ and $j = 1, \dots 43$.

```{r colSetup}
#| echo: false

blue_pal <- paletteer::paletteer_c("ggthemes::Classic Blue", 98)
orange_pal <- paletteer::paletteer_c("ggthemes::Classic Orange", 98)
gray_pal = paletteer::paletteer_c("ggthemes::Classic Gray", 98)
viridis_pal <- paletteer::paletteer_c("viridis::viridis", 98)
```

### Dynamic GNAR Fit

```{r dynGNAR}

bru_struct_GNARfit <- temp_trend_cv_GNARfit(longitudinal_network_data = 100.0 * MXD2, network = bur_netGNAR, alpha_order = alpha_order, beta_order = beta_order, weight_matrix = W_bru_data, 
                                         covariate_groups = bru_data_community_indicators, interaction_groups = bru_data_interactions, rmax = rmax, iter = 100)

bru_res_interaction_fit <- community_interaction_gnar_fit(bru_struct_GNARfit$`GNAR residuals`, bur_netGNAR, alpha_order, beta_order, W_bru_data, bru_data_community_indicators, 
                                                          bru_data_interactions)
summary(bru_res_interaction_fit)
#print(bru_struct_GNARfit$`GNAR parameters`)
print(sum(bru_struct_GNARfit$`GNAR residuals`[2:44, ]^2))
print(sum(bru_struct_GNARfit$`Residuals Vector`^2) / (98 * 43 - 34))
print(sum(abs(bru_struct_GNARfit$`Residuals Vector`)) / (98 * 43 - 34))
print(sqrt(sum(bru_struct_GNARfit$`GNAR parameters`)^2))
bru_res_interaction_fit_ts <- bru_struct_GNARfit$`Residuals Matrix`


plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(bru_res_interaction_fit_ts[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), bru_res_interaction_fit_ts[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), bru_res_interaction_fit_ts[, j])))
}
l1_dgnar = l1_norm
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))

hist(bru_struct_GNARfit$`Residuals Vector`, breaks = 50, xlab = expression(hat(u)), main = "Dynamic GNAR Fit Residuals", cex = 5)

print(GNAR_AIC(bru_struct_GNARfit$`Residuals Vector`, 98, 90))
GNAR_BIC(bru_struct_GNARfit$`Residuals Vector`, 98, 90)
GNAR_HQC(bru_struct_GNARfit$`Residuals Vector`, 98, 90)
nacf_vals <- corbit_plot_cpp(bru_struct_GNARfit$`Residuals Matrix`, bur_netGNAR, 10, 4, W_bru_data)
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))
nacf_dgnar = max(abs(nacf_vals))
```

### Global-$\alpha$ GNAR Fit

```{r gloGNAR}
bru_global_fit <- global_gnar_fit(bru_vts, bur_netGNAR, 1, 3, W_bru_data)
summary(bru_global_fit)
bru_global_fit_ts = residuals_matrix(bru_global_fit$residuals, 98)
print(sum((bru_global_fit$residuals)^2) / (98 * 43 - 4))
print(sum(abs(bru_global_fit$residuals)))
print(sqrt(sum((bru_global_fit$coefficients)^2)))
print(sum(abs(bru_global_fit$coefficients)))

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(bru_global_fit_ts[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), bru_global_fit_ts[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), bru_global_fit_ts[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_gnar = l1_norm

hist(bru_global_fit$residuals, breaks = 50, xlab = expression(hat(u)), main = "Global-alpha GNAR Fit Residuals", cex = 5)
GNAR_AIC(bru_global_fit$residuals, 98, 3)
GNAR_BIC(bru_global_fit$residuals, 98, 3)
GNAR_HQC(bru_global_fit$residuals, 98, 3)
nacf_vals <- corbit_plot_cpp(residuals_matrix(bru_global_fit$residuals, 98), bur_netGNAR, 10, 4, W_bru_data)
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))
nacf_gnar = max(abs(nacf_vals))

```

### Community-$\alpha$ with $\gamma$-Interactions GNAR Fit

```{r comGNAR}
bru_interaction_fit <- community_interaction_gnar_fit(bru_vts, bur_netGNAR, alpha_order, beta_order, W_bru_data, bru_data_community_indicators, bru_data_interactions)
summary(bru_interaction_fit)

bru_interaction_fit_ts = residuals_matrix(bru_interaction_fit$residuals, 98)
sum((bru_interaction_fit$residuals)^2) / (43 * 98 - 34)
sum(abs(bru_interaction_fit$residuals))
sqrt(sum((bru_interaction_fit$coefficients)^2))
sum(abs(bru_interaction_fit$coefficients))

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(bru_interaction_fit_ts[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), bru_interaction_fit_ts[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), bru_interaction_fit_ts[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_intgnar = l1_norm

hist(bru_interaction_fit$residuals, breaks = 50, xlab = expression(hat(u)), main = "Interactions GNAR Fit Residuals", cex = 5)
GNAR_AIC(bru_interaction_fit$residuals, 98, 0)
GNAR_BIC(bru_interaction_fit$residuals, 98, 30)
GNAR_HQC(bru_interaction_fit$residuals, 98, 30)
nacf_vals <- corbit_plot_cpp(residuals_matrix(bru_interaction_fit$residuals, 98), bur_netGNAR, 10, 4, W_bru_data)
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))
nacf_intgnar = max(abs(nacf_vals))
```

### Smoother Fit: 98 Independent Splines

```{r splines}
mu_hat_mat <- temporal_trend_spline_smoother_cv(spline_residuals = bru_vts[2:44, ], vts_dimension = 98, plag = 1)
mu_res_mat <- bru_vts[2:44, ] - mu_hat_mat[2:44, ]
mu_hat_res <- recursive_dmat_cols_cpp(lapply(1:43, function(x) {return(as.matrix(mu_res_mat[x, ]))}))

sum(mu_hat_res^2) / (98 * 43 - 1)
sum(abs(mu_hat_res)) / (98 * 43 - 1)

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(mu_res_mat[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), mu_res_mat[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), mu_res_mat[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_splines = l1_norm

hist(mu_hat_res, breaks = 50, xlab = expression(hat(u)), main = "98 Smoothers Fit Residuals", cex = 5)
GNAR_AIC(mu_hat_res, 98, 0)
GNAR_BIC(mu_hat_res, 98, 30)
GNAR_HQC(mu_hat_res, 98, 30)
nacf_vals <- corbit_plot_cpp(mu_res_mat, bur_netGNAR, 10, 4, W_bru_data)
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))
nacf_splines = max(abs(nacf_vals))
```

### Sparse VAR Fit

```{r spVAR}
sparse_var_fit <- fitVAR(100 * MXD2, p = 1, nlambda = 100)
theta_mat = sparse_var_fit$A[[1]]


sum(abs(sparse_var_fit$residuals))
sparse_var_fit$lambda
sparse_var_res <- sparse_var_fit$residuals[2:44, ]

sqrt(sum(theta_mat^2))
b0 = theta_mat
b0[abs(b0) > 0.0] = 1
cat(paste0("Number of non-zero coefficients: ", sum(rowSums(b0))))
sum((sparse_var_res)^2) / (98 * 43 - sum(rowSums(b0)))

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(sparse_var_res[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), sparse_var_res[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), sparse_var_res[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_spvar = l1_norm

hist(recursive_dmat_cols_cpp(lapply(1:98, function(j) {return(as.matrix(sparse_var_fit$residuals[2:44, j]))})), breaks = 50, xlab = 'u', main ='Sparse VAR Fit Res', cex = 5)

GNAR_AIC(recursive_dmat_cols_cpp(lapply(1:98, function(j) {return(as.matrix(sparse_var_fit$residuals[2:44, j]))})), 98, sum(rowSums(b0)))
GNAR_BIC(recursive_dmat_cols_cpp(lapply(1:98, function(j) {return(as.matrix(sparse_var_fit$residuals[2:44, j]))})), 98, sum(rowSums(b0)))
GNAR_HQC(recursive_dmat_cols_cpp(lapply(1:98, function(j) {return(as.matrix(sparse_var_fit$residuals[2:44, j]))})), 98, sum(rowSums(b0)))
nacf_vals <- corbit_plot_cpp(sparse_var_fit$residuals[2:44, ], bur_netGNAR, 10, 4, W_bru_data)
max(abs(nacf_vals))
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))
VAR_mat_plot(1, 1, theta_mat)

nacf_var = max(abs(nacf_vals))
```

### Null Model

```{r null}
naive_mean <- bru_vts[2:44, ]
naive_mean_vec <- recursive_dmat_cols_cpp(lapply(2:44, function(x) {return(as.matrix(bru_vts[x, ]))}))
sum(naive_mean_vec^2) / (98 * 43 - 1)
sum(abs(naive_mean_vec)) / (98 * 43 - 1)

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(naive_mean[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), naive_mean[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), naive_mean[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_null = l1_norm

hist(recursive_dmat_cols_cpp(lapply(1:43, function(x) {return(as.matrix(naive_mean[x, ]))})), breaks = 50, xlab = expression(hat(u)), main = "Centred Data", cex = 5)

nacf_vals <- corbit_plot_cpp(bru_vts, bur_netGNAR, 10, 4, W_bru_data)
max(abs(nacf_vals))
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))

nacf_null = max(abs(nacf_vals))
```

### Naive Model

```{r naive}
naive_res <- 100 * t(vapply(2:44, function(x) {MXD2[x, 1:98] - MXD2[x - 1, 1:98]}, rep(0, 98)))
naive_res_vec <- recursive_dmat_cols_cpp(lapply(1:43, function(x) {return(as.matrix(naive_res[x, ]))}))
sum((naive_res)^2) / (98 * 43 - 1)
sum(abs(naive_res)) / (98 * 43 - 1)

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(naive_res[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), naive_res[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), naive_res[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_naive = l1_norm

hist(recursive_dmat_cols_cpp(lapply(1:43, function(x) {return(as.matrix(naive_res[x, ]))})), breaks = 50, xlab = expression(hat(u)), main = "Naive Fit Residuals", cex = 5)

nacf_vals <- corbit_plot_cpp(naive_res, bur_netGNAR, 10, 4, W_bru_data)
max(abs(nacf_vals))
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))
nacf_naive = max(abs(nacf_vals))
```

### 98 ARs Fits

```{r arFit}

ar_residuals <- vapply(1:98, function(x){forecast::auto.arima(bru_vts[1:44, x], d = 0, D = 0, max.p = 1, max.q = 0,
                      max.P = 0, max.Q = 0, stationary = TRUE, seasonal = FALSE, ic = 'bic',
                      allowmean = FALSE, allowdrift = FALSE, trace = FALSE)$residuals}, rep(0, 44))[2:44, ]

ar_coeffs <- vapply(1:98, function(x){forecast::auto.arima(bru_vts[1:44, x], d = 0, D = 0, max.p = 1, max.q = 0, max.P = 0, max.Q = 0, stationary = TRUE, seasonal = FALSE, ic = 'bic',
                      allowmean = FALSE, allowdrift = FALSE, trace = FALSE)$coef[1]}, 1)

ar_coeffs[is.na(ar_coeffs)] = 1

ar_res_vec <- recursive_dmat_cols_cpp(lapply(1:43, function(x) {return(as.matrix(ar_residuals[x, ]))}))
sum(ar_res_vec^2) / (98 * 42)
sum(abs(ar_res_vec)) / (98 * 42)

plot(rep(0, 43), type = 'l', col = 'blue', ylim = c(-2, 2), ylab = expression(hat(u)), xlab = 't', cex = 5)
l1_norm = 0
for (j in 1:98) {
  points(c(1:43), c(ar_residuals[, j]), pch = 16)
  lines(c(1:43), vapply(1:43, function(i) {c(1, i) %*% least_squares_solver(cbind(rep(1, 43), c(1:43)), naive_res[, j])}, 0.0), col = orange_pal[j],
        lty = 2)
  l1_norm = l1_norm + sum(abs(least_squares_solver(cbind(rep(1, 43), c(1:43)), ar_residuals[, j])))
}
cat(paste0("Sum of residual linear trend coefficients: ", as.character(l1_norm)))
l1_ar = l1_norm

hist(recursive_dmat_cols_cpp(lapply(1:43, function(x) {return(as.matrix(ar_residuals[x, ]))})), breaks = 50, xlab = expression(hat(u)), main = "98 ARs Fit Residuals", cex = 5)

nacf_vals <- corbit_plot_cpp(ar_residuals, bur_netGNAR, 10, 4, W_bru_data)
max(abs(nacf_vals))
cat(paste0("Maximum NACF Value: ", max(abs(nacf_vals))))

nacf_ar = max(abs(nacf_vals))
```

## Model Comparisons

The table below reproduces Table 1 and summarizes model comparisons.

```{r comparisonTable}
#| echo: false

mod_comps <- matrix(rep(0, 8 * 5), nrow = 8, ncol = 5)
rownames(mod_comps) <- c("Dynamic GNAR", "Global-a GNAR", "g-Interactions GNAR", "98 Smoothers", "Sparse VAR", "98 AR", "Naive Model", "Null Model")
colnames(mod_comps) <- c("MSE", "MAE", "NACF", "Est. Norm", "Res. Trend")

mod_comps[1, ] <- c(sum(bru_struct_GNARfit$`Residuals Vector`^2) / 43, sum(abs(bru_struct_GNARfit$`Residuals Vector`)) / 43, nacf_dgnar, sqrt(sum(bru_struct_GNARfit$`GNAR parameters`)^2), l1_dgnar)

mod_comps[2, ] <- c(sum(bru_global_fit$residuals^2) /43, sum(abs(bru_global_fit$residuals)) /43, nacf_gnar, sqrt(sum(bru_global_fit$coefficients)^2), l1_gnar)

mod_comps[3, ] <- c(sum(bru_interaction_fit$residuals^2) /43, sum(abs(bru_interaction_fit$residuals)) /43, nacf_intgnar, sqrt(sum(bru_interaction_fit$coefficients)^2), l1_intgnar)


mod_comps[4, ] <- c(sum(mu_hat_res^2) /43, sum(abs(mu_hat_res)) /43, nacf_splines, 0, l1_splines)

mod_comps[5, ] <- c(sum((sparse_var_res)^2) / 43, sum(abs(sparse_var_res)) / 43, nacf_var, sqrt(sum(theta_mat^2)), l1_spvar)

mod_comps[6, ] <- c(sum(ar_res_vec^2) / 43, sum(abs(ar_res_vec)) / 43, nacf_ar, sqrt(sum(ar_coeffs^2)), l1_ar)

mod_comps[7, ] <- c(sum(naive_res^2) / 43, sum(abs(naive_res)) / 43, nacf_naive, 0, l1_naive)

mod_comps[8, ] <- c(sum(naive_mean_vec^2) /43, sum(abs(naive_mean_vec)) / 43, nacf_null, 0, l1_null)

#mod_comps[, 1] <- 100.0 * mod_comps[, 1]

#mod_comps[, 2] <- 100.0 * mod_comps[, 2]

mod_comps_frame <- data.frame(mod_comps, check.names = FALSE)

knitr::kable(mod_comps_frame, format = 'html', digits = c(4, 4, 2, 2, 2), col.names = c("MSE", "MAE", "NACF", "Est. Theta Norm", "Res. Trend Coeff."))

#knitr::kable(mod_comps_frame, format = 'latex', digits = c(2, 2, 2, 2, 2), col.names = c("MSE", "MAE", "NACF", "Est. Norm", "Res. Trend"))
```

## Temporal Trend

We plot the fitted temporal trend, which also accounts for autocorrelation in the data. The plot shows the main effect of this group-wise decreasing trends.

```{r longitudinalPlots}
#| echo: false
#| message: false

loc1_sat1_nts <- t(vapply(seq(1:44), function(x) {return(loc1_sat1 * MXD2[x, ])}, rep(0.0, 98)))
loc1_sat2_nts <- t(vapply(seq(1:44), function(x) {return(loc1_sat2 * MXD2[x, ])}, rep(0.0, 98)))
loc2_sat1_nts <- t(vapply(seq(1:44), function(x) {return(loc2_sat1 * MXD2[x, ])}, rep(0.0, 98)))
loc2_sat2_nts <- t(vapply(seq(1:44), function(x) {return(loc2_sat2 * MXD2[x, ])}, rep(0.0, 98)))
loc3_sat1_nts <- t(vapply(seq(1:44), function(x) {return(loc3_sat1 * MXD2[x, ])}, rep(0.0, 98)))
loc3_sat2_nts <- t(vapply(seq(1:44), function(x) {return(loc3_sat2 * MXD2[x, ])}, rep(0.0, 98)))

loc1_sat1_trend <- bru_struct_GNARfit$`temporal trend fitted values`[, 1:10]
loc1_sat2_trend <- bru_struct_GNARfit$`temporal trend fitted values`[, 2:20]
loc2_sat1_trend <- bru_struct_GNARfit$`temporal trend fitted values`[, 21:37]
loc2_sat2_trend <- bru_struct_GNARfit$`temporal trend fitted values`[, 72:86]
loc3_sat1_trend <- bru_struct_GNARfit$`temporal trend fitted values`[, 38:60]
loc3_sat2_trend <- bru_struct_GNARfit$`temporal trend fitted values`[, c(61:71, 87:98)]



#########
mean_loc1_sat1_trend <- rowMeans(loc1_sat1_trend)
loc1_sat1_quantile <- vapply(1:44, function(x) {quantile(loc1_sat1_trend[x, ], probs = c(0.0, 1))}, rep(0.0, 2))
plot(mean_loc1_sat1_trend[1:44], type ='l', ylim = c(-1, 1), col = 'forestgreen', lwd=3, pch = 16,
     main = "Location 1 Satellite 1", xlab = '', cex.main = 2, cex.lab = 2, ylab = "Trend Estimate" )+
  lines(c(1:44), c(loc1_sat1_quantile[1, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  lines(c(1:44), c(loc1_sat1_quantile[2, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  for (j in 1:10) {
    #lines(c(1:44), -.63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)+
    #lines(c(1:44), .63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)+
    points(c(1:44), 100.0 * (loc1_sat1_nts[1:44, j] - mean(loc1_sat1_nts[1:44, j])),
           col = 'black', pch = 16, cex = 1)
    lines(c(1:44), bru_struct_GNARfit$`temporal trend fitted values`[1: 44, j], col = viridis_pal[j], lwd = 1, type = 'l', lty = 1)
  }
############################
mean_loc2_sat1_trend <- rowMeans(loc2_sat1_trend)
loc2_sat1_quantile <- vapply(1:44, function(x) {quantile(loc2_sat1_trend[x, ], probs = c(0.0, 1))}, rep(0.0, 2))
plot(mean_loc2_sat1_trend[1:44], type ='l', ylim = c(-1.5, 1), col = 'forestgreen', lwd=3, pch = 16,
     main = "Location 2 Satellite 1", ylab = expression(hat(mu)["t"]), xlab = "", cex.main = 2, cex.lab = 2 )+
  lines(c(1:44), c(loc2_sat1_quantile[1, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  lines(c(1:44), c(loc2_sat1_quantile[2, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  for (j in 21:37) {
    points(c(1:44), 100.0 * (loc2_sat1_nts[1:44, j] - mean(loc2_sat1_nts[1:44, j])),
           col = 'black', pch = 16, cex = 1)
    lines(c(1:44), bru_struct_GNARfit$`temporal trend fitted values`[1: 44, j], col = viridis_pal[j], lwd = 1, type = 'l', lty = 1)
    #lines(c(1:44), -.63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    #lines(c(1:44), .63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
  }
########
mean_loc3_sat1_trend <- rowMeans(loc3_sat1_trend)
loc3_sat1_quantile <- vapply(1:44, function(x) {quantile(loc3_sat1_trend[x, ], probs = c(0.0, 1))}, rep(0.0, 2))
plot(mean_loc3_sat1_trend[1:44], type ='l', ylim = c(-1.5, 1.5), col = 'forestgreen', lwd=3, pch = 16,
     main = "Location 3 Satellite 1", ylab = expression(hat(mu)["t"]), xlab = "", cex.main = 2, cex.lab = 2 )+
  lines(c(1:44), c(loc3_sat1_quantile[1, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  lines(c(1:44), c(loc3_sat1_quantile[2, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  for (j in 38:60) {
    points(c(1:44), 100.0 * (loc3_sat1_nts[1:44, j] - mean(loc3_sat1_nts[1:44, j])),
           col = 'black', pch = 16, cex = 1)
    lines(c(1:44), bru_struct_GNARfit$`temporal trend fitted values`[1:44, j], col = viridis_pal[j], lwd = 1, type = 'l', lty = 1)
    #lines(c(1:44), -.63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    #lines(c(1:44), .63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
  }
################
mean_loc1_sat2_trend <- rowMeans(loc1_sat2_trend)
loc1_sat2_quantile <- vapply(1:44, function(x) {quantile(loc1_sat2_trend[x, ], probs = c(0.0, 1))}, rep(0.0, 2))
plot(mean_loc1_sat2_trend[1:44], type ='l', ylim = c(-1.5, 1), col = 'forestgreen', lwd=3, pch = 16,
     main = "Location 1 Satellite 2", ylab = "Trend Estimate", xlab = "Time: t", cex.main = 2, cex.lab = 2 )+
  lines(c(1:44), c(loc1_sat2_quantile[1, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  lines(c(1:44), c(loc1_sat2_quantile[2, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  for (j in 11:20) {
    #lines(c(1:44), -.63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    #lines(c(1:44), .63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    points(c(1:44), 100.0 * (loc1_sat2_nts[1:44, j] - mean(loc1_sat2_nts[1:44, j])),
           col = 'black', pch = 16, cex = 1)
    lines(c(1:44), bru_struct_GNARfit$`temporal trend fitted values`[1: 44, j], col = viridis_pal[j], lwd = 1, type = 'l', lty = 1)
  }
########
mean_loc2_sat2_trend <- rowMeans(loc2_sat2_trend)
loc2_sat2_quantile <- vapply(1:44, function(x) {quantile(loc2_sat2_trend[x, ], probs = c(0.0, 1))}, rep(0.0, 2))
plot(mean_loc2_sat2_trend[1:44], type ='l', ylim = c(-2, 2.5), col = 'forestgreen', lwd=3, pch = 16,
     main = "Location 2 Satellite 2", ylab = expression(hat(mu)["t"]), xlab = "Time: t", cex.main = 2, cex.lab = 2 )+
  lines(c(1:44), c(loc2_sat2_quantile[1, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  lines(c(1:44), c(loc2_sat2_quantile[2, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  for (j in 72:86) {
    points(c(1:44), 100.0 * (loc2_sat2_nts[1:44, j] - mean(loc2_sat2_nts[1:44, j])),
           col = 'black', pch = 16, cex = 1)
    lines(c(1:44), bru_struct_GNARfit$`temporal trend fitted values`[1: 44, j], col = viridis_pal[j], lwd = 1, type = 'l', lty = 1)
    #lines(c(1:44), -.63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    #lines(c(1:44), .63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
  }
#############
mean_loc3_sat2_trend <- rowMeans(loc3_sat2_trend)
loc3_sat2_quantile <- vapply(1:44, function(x) {quantile(loc3_sat2_trend[x, ], probs = c(0.0, 1))}, rep(0.0, 2))
plot(mean_loc3_sat2_trend[1:44], type ='l', ylim = c(-1.5, 1.5), col = 'forestgreen', lwd=3, pch = 16,
     main = "Location 3 Satellite 2", ylab = expression(hat(mu["t"])), xlab = "Time: t" , cex.main = 2, cex.lab = 2)+
  lines(c(1:44), c(loc3_sat2_quantile[1, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  lines(c(1:44), c(loc3_sat2_quantile[2, 1:44]), col = 'blue', lwd = 2, type = 'l', lty = 2)+
  for (j in c(61:71, 87:98)) {
    #lines(c(1:44), -.63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    #lines(c(1:44), .63 + bru_struct_GNARfit$`temporal trend fitted values`[1:44, j] + 
    #        bru_struct_GNARfit$`GNAR residuals`[1:44, j], col = 'orange', type = 'h', lwd = 1)
    points(c(1:44), 100.0 * (loc3_sat2_nts[1:44, j] - mean(loc3_sat2_nts[1:44, j])), col = 'black', pch = 16, cex = 1)
    lines(c(1:44), bru_struct_GNARfit$`temporal trend fitted values`[1: 44, j], col = viridis_pal[j], lwd = 1, type = 'l', lty = 1)
  }


```
